---
title: "Heart-Disease Directory"
author: "Andres Gomez"
date: "2023-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install tidyverse package required to run the program for end user
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")

# Install knitr package required to run the program for end user
if(!require(knitr)) install.packages("knitr", repos = "http://cran.us.r-project.org")

# Install sjstats package require to run the program for end user
if(!require(sjstats)) install.packages("sjstats", repos = "http://cran.us.r-project.org")

if(!require(cowplot)) install.packages("cowplot", repos = "http://cran.us.r-project.org") 
#this is for the plot_grid function to show the scatterplots in part one

# Obtaining the data:

# Retrieve the Heart Disease data from GitHub
url <- 'https://raw.githubusercontent.com/jholland5/COMP4299/main/heartData.csv'

# Read the data into R and name it concrete.
heartData <- read_csv(url)
```

# Heart Disease Database Analysis

## Part 1 - Heart Disease Data Analysis

The data presented in this report contains 303 rows of 10 numeric variables relating to heart disease. The intent of the data in to predict the presence of a heart disease in a patient. Said information is represented with an integer value from 0 (no presence) to 1 (present). The variables are listed below along with the corresponding variable name in parenthesis, and consist of numeric values.

##### Variable names and descriptions

* **Age** (age) describes the age in years.
* **Sex** (sex) describes the sex of the patient in values of 1 (male) and 0 (female).

* **Resting blood pressure** (sys_press) value of the resting blood pressure in $\frac{mm}{Hg}$ on admission to the hospital.

* **Serum Cholesterol** (chol) value of the serum cholesterol in $\frac{mg}{dl}$.

* **Fasting Blood Sugar** (fast_bsug) fasting blood sugar in terms of: fasting blood sugar > 120 $\frac{mg}{dl}$, whether true (1) or false (0). 

* **Resting electrocardiographic results** (recg) values represent: **0** = Normal, **1** = having ST-T wave abnormality (T wave inversions and/or ST elevation or depression of > 0.05 mV), **2** = showing probable or definite left ventricular hypertrophy by Estes' criteria.

* **Maximum heart rate** (max_hrate) value of maximum heart rate achieved.

* **Old Peak** (opk) ST depression induced by exercise relative to rest.

* **Number of Vessels** (nvess) Number of major vessels in values ranged from 0 to 3.

* **Level** (level) is the value to predict. Values represent absence (0) or presence (1)  of heart disease. This is the attribute to predict.

<br>

## Part 2 - Preparing the Data and Analyzing the Predictors:

##### The following is a glimpse at the first 6 rows of the data
```{r, echo=FALSE}
#Note: The first two values are not shown as output.
#Transforming the data into a data frame:
heart_DataFrame <- as.data.frame(heartData)

#Inserting the first 6 rows of data into a variable:
table_heartData <- heart_DataFrame[1:6, 1:10]

#Displaying the variable on a table.
knitr::kable(table_heartData)
```

<br>

##### Summary Statistics of the heart disease data:

```{r, echo=FALSE}
#Inserting a summary of the first three variables into another variable: 
heartDataSummary1 <- summary(heart_DataFrame[,1:3])
#running a summary on them and printing it in a table
knitr::kable(heartDataSummary1)

#Inserting a summary of the second three variables into another variable: 
heartDataSummary2 <- summary(heart_DataFrame[,4:6])
#running a summary on them and printing it in a table
knitr::kable(heartDataSummary2)

#Inserting a summary of the third three variables into another variable: 
heartDataSummary3 <- summary(heart_DataFrame[,7:9])
#running a summary on them and printing it in a table
knitr::kable(heartDataSummary3)
```

<br>

##### Vizualization of the heart disease data with histograms:

In this subsection, the report will display faceted histogram plots of the different variables in relation with the predictor. The faceted histogram plots are displayed below.

```{r, echo = FALSE}
#Setting the data into Training and testing set:
#We are getting rid of the attribute to predict (level), and those values that contain 0 and 1.
heartTrain <- heartData %>% select(-c(sex, fast_bsug))
heartTest <- heartData %>% select(-c(sex, fast_bsug))

#Scaling the predictors to remove the bias:
heartTrain[,1:6] <- as.data.frame(scale(heartTrain[,1:6]), show_col_types = FALSE)
heartTest[,1:6] <- as.data.frame(scale(heartTest[,1:6]), show_col_types = FALSE)

#Faceted Histogram:
heartTrain %>%
  gather(-level, key = "var", value = "value") %>% 
  ggplot(aes(x = value, color = "blue")) +
  geom_histogram(bins=18,color = "black",fill = "blue") +
  facet_wrap(~ var, scales = "free") + 
  labs(title = "Distributions of Predictors") + 
  theme(axis.title.x=element_blank(),axis.title.y = element_blank())
```

The report will now procede to determine the best single predictor.

<br>

## Part 3 - Visualizing a Logistic Model with One Predictor:

##### Determining the best predictor:

The following visualizations illustrate two predictors at the same time to prove whether they separate the response variable well, or not.

```{r, echo=FALSE}
splot <- heartTrain %>% ggplot() 
splot + geom_point(aes(max_hrate,sys_press,col = as.factor(level))) +
  stat_ellipse(aes(max_hrate,sys_press,col=as.factor(level)),
               type = "norm",lwd =1.5,level = .7) +
  scale_color_manual(values = c("red", "blue")) +
  labs(col = "1 = level", title = "Maximum Heart Rate (max_hrate) vs Resting Blood Pressure (sys_press)") 
```

By looking at the graphs, the report can determine that the possible best predictors were **max_hrate** and **sys_press**, however, it is possible to observe that **max_hrate** separates the levels of heart disease absent (0) or present (1). 

The report will present another perspective by using **age** instead of **sys_press**.

```{r, echo=FALSE}
splot <- heartTrain %>% ggplot() 
splot + geom_point(aes(max_hrate,age,col = as.factor(level))) +
  stat_ellipse(aes(max_hrate,age,col=as.factor(level)),
               type = "norm",lwd =1.5,level = .7) +
  scale_color_manual(values = c("red", "blue")) +
  labs(col = "1 = level", title = "Maximum Heart Rate (max_hrate) vs Resting Blood Pressure (sys_press)") 
```

Once again, it is possible to observe the importance of **max_hrate**.

<br>

##### Logistic Model with One Predictor:

The report will now display a visualization of the results accompanied by the intercept and predictor.
```{r, echo=FALSE}
#Producing logistics Model
glm_fit <- heartTrain %>% 
  glm(level ~ max_hrate,data =.,family = "binomial") # Logistic model
p_hat_glm <- predict(glm_fit,heartTest, type = "response") # probabilities
y_hat_glm <- factor(ifelse(p_hat_glm > .5,1,0),levels = c("0","1")) #0-1

glm_fit$coefficients

t <- heartTest$max_hrate      # inputs to the model predictions
y <- 1/(1+exp(-0.1622701)*exp(-0.9946601*t))   # probabilities
heartTest$Predictor <- as.factor(ifelse(heartTest$level == 1,"Present","Absent")) # Labels
heartTest$y <- y
vis <- heartTest %>% ggplot()
vis + geom_jitter(aes(max_hrate,y,col=Predictor),size = 3,width = .01,height = .02) +
  geom_abline(slope = 0, intercept = .5,col = "red") +
  scale_color_manual(values = c("Blue", "Red")) +
  labs(title = "Visualizing Logistic Regression with One Predictor",
       subtitle = "Accuracy is 24/32 = .75")

T <- table(heartTest$level,y_hat_glm)
T
```
<br>

## Conclusion:

It is possible to now understand that the **max_hrate** variable may not be enough to understand whether a heart disease is present or absent due to several misclassifications that can be observed in the previous Logistics Model.
